---
  title: "Practical Machine Learning Course Project"
author: "Shivansh Shrivastava"
date: "October 21, 2020"
output: html_document
---
  
  
  
  # Project Description
  
  > **Background**
  > 
  > Using devices such as *Jawbone Up*, *Nike FuelBand*, and *Fitbit* it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement -- a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 


# My Code

```{r, include=FALSE}
setwd("C:/Users/Jeff/Coursera/Machine Learning/Course_Project_GitHub/practicalmachinelearning")
rm(list=ls())
```

```{r, message=FALSE}
# load necessary libraries
library(caret)
library(randomForest)
```

```{r}
# read in the data and identify the NA's
traindata <- read.csv("pml-training.csv", na.strings = c("NA", "#DIV/0!", ""))
testdata  <- read.csv("pml-testing.csv", na.strings = c("NA", "#DIV/0!", ""))
# remove NA columns for the training and testing data
comps <- complete.cases(t(traindata)) & complete.cases(t(testdata))
traindata <- traindata[,comps]
testdata  <- testdata[,comps]
# remove columns with data that isn't useful
traindata <- traindata[,-c(1,3,4,5,6,7)]
testdata <- testdata[,-c(1,3,4,5,6,7)]
# data splitting
set.seed(111)
inTrain <- createDataPartition(traindata$classe, p=0.6, list=FALSE)
traindata2 <- traindata[inTrain,]
validation <- traindata[-inTrain,]
# fit a model
modFit <- randomForest(classe~., data=traindata2)
# the results on the training set
trainresults <- predict(modFit, traindata2)
trainacc <- sum(trainresults==traindata2$classe)/length(trainresults)
paste("Accuracy on training set =",trainacc)
# the results on the validation set
validresults <- predict(modFit, newdata=validation)
validacc <- sum(validresults==validation$classe)/length(validresults)
paste("Accuracy on validation set =",validacc)
# the results on the test set
testresults <- predict(modFit, newdata=testdata)
print("Classifications on the test set:"); testresults
```